---
title: "Seamus Gallivan Crime Project"
author: "Seamus Gallivan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook: default
---

the tidyverse, dplyr, stringr, car and modeest package will be used for this R markdown file.

```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(modeest)
library(car)
```

Compile date and session info

```{r}
# print and get runtime ----
date()

# session info ----
sessionInfo()
```

read in the nola_crime_2018 dataset.

```{r}
nola_crime <- read_csv(file="/Users/seamusgallivan/Desktop/nola_crime_2018.csv")
```

**How many observations and variables in the dataset?**

```{r}
str(nola_crime)
```
*Observations = 9970, Variables = 24*

**How many unique item numbers are in the dataset?**

```{r}
length(unique(nola_crime$Item_Number))
```
*9384 unique item numbers*

**How many aggravated crimes were reports in the Irish Channel?**

```{r}
nola_crime %>% 
  select(GNOCDC_LAB, Signal_Description, Item_Number) %>%
  filter(str_detect(Signal_Description, "AGG")) %>%
filter(GNOCDC_LAB=="IRISH CHANNEL")
```
*1 Aggravated Crime in The Irish Channel Neighorhood*

**Which 2 neighborhoods featured the most crime reports?**

```{r}
nola_crime <- nola_crime %>% 
  group_by(GNOCDC_LAB) %>%
  mutate(unique_item_numbers = n_distinct(Item_Number)) 
nola_crime %>%
  arrange(desc(unique_item_numbers)) %>%
  select(GNOCDC_LAB, unique_item_numbers) %>%
  distinct(GNOCDC_LAB, .keep_all = TRUE) 
```
*Two largest number of crime reports = FRENCH QUARTER & LITTLE WOODS*

**Calculate the rate of crime per 1,000 persons in each neighborhood.**

```{r}
nola_crime <- nola_crime %>% mutate(
  crime_rate = (unique_item_numbers / pop) * 1000
)
nola_crime %>%
  select(pop,unique_item_numbers,crime_rate,GNOCDC_LAB) %>%
  distinct(GNOCDC_LAB, .keep_all = TRUE) 
nola_crime %>%
  group_by(GNOCDC_LAB) %>%
  select(GNOCDC_LAB,crime_rate) %>%
  distinct(GNOCDC_LAB, crime_rate)
```

**What is the largest neighborhood by population in New Orleans?**

```{r}
nola_crime %>%
  select(GNOCDC_LAB, pop) %>%
  arrange(desc(pop))
```
*Largest neighborhood by population = LITTLE WOODS, population = 44947*

**Which neighborhood had the lowest crime rate per 1,000 persons?**

```{r}
nola_crime %>%
  select(pop, unique_item_numbers, GNOCDC_LAB, crime_rate) %>%
  arrange(crime_rate)
```
*Lowest crime rate = LAKEWOOD, population = 13989, unique crimes = 10*

**Calculate the rate ratio for each neighborhood for all unique crimes treating LAKEWOOD as the reference.**

```{r}
nola_crime %>%
  filter(GNOCDC_LAB=="LAKEWOOD")
nola_crime <- nola_crime %>% 
  mutate(
    rate_ratio = crime_rate / 0.7148474
  )
nola_crime %>%
  select(GNOCDC_LAB, rate_ratio) %>%
  distinct(GNOCDC_LAB, rate_ratio)
```

**Calculate the rate ratio for each neighborhood for all unique aggravated crimes using FRENCH QUARTER as the reference.**

```{r}
nola_crime1 <- nola_crime %>%
  group_by(GNOCDC_LAB) %>%
  filter(str_detect(Signal_Description, "AGG")) %>%
  mutate(
    unique_agg_crimes = n_distinct(Item_Number)
  ) %>%
  mutate(
    unique_agg_crime_rate = (unique_agg_crimes / pop) * 1000
  ) 
nola_crime1 %>%
  group_by(GNOCDC_LAB) %>%
  select(GNOCDC_LAB, unique_agg_crime_rate) %>%
  filter(str_detect(GNOCDC_LAB, "FRENCH QUARTER"))
nola_crime1 <- nola_crime1 %>% 
  group_by(GNOCDC_LAB) %>%
  mutate(
    agg_rate_ratio = unique_agg_crime_rate / 0.9886633
  )
nola_crime1 %>%
  group_by(GNOCDC_LAB)%>%
  select(agg_rate_ratio, GNOCDC_LAB) %>%
  distinct(agg_rate_ratio, GNOCDC_LAB)
```

**What was the most common cause for a police report and what was the most common charge filed?**

```{r}
mfv(nola_crime$Signal_Description)
mfv(nola_crime$Charge_Description, na_rm = TRUE)
```
*most common cause for police report = "DISTURBANCE (DOMESTIC)"*
*most common charge that is filed = "THEFT"*

**Calculate the rate of domestic disturbance by neighborhood.**

```{r}
nola_crime1 <- nola_crime %>%
  filter(Signal_Description == "DISTURBANCE (DOMESTIC)") %>%
  mutate(
    domestic_disturbances = n_distinct(Item_Number)
  )
nola_crime1 <- nola_crime1 %>%
  mutate(
    dom_dist_rate = (domestic_disturbances / pop) * 1000
  )
nola_crime1 %>%
  select(GNOCDC_LAB, dom_dist_rate) %>%
  distinct(GNOCDC_LAB, dom_dist_rate)
```
**Regression model.Predictor: mean victim age by neighborhood. Response: crime rate by neighborhood**
```{r}
nola_crime2 <- nola_crime %>%
  group_by(GNOCDC_LAB,crime_rate) %>% 
  select(GNOCDC_LAB,crime_rate,Victim_Age) %>%
  summarize(m = mean(Victim_Age, na.rm = TRUE))
nola_crime2
  linearreg <- lm(m ~ crime_rate, data = nola_crime2)
  print(linearreg)
  summary(linearreg)
```

**Which neighborhood has the worst police responsiveness?**

```{r}
nola_crime %>%
  select(GNOCDC_LAB, crime_rate) %>%
  arrange(crime_rate)
```
*Since LAKEWOOD had the lowest crime rate, it also had the lowest police reponsiveness. LAKEWOOD had the least value of least amonunt of Signals per population*

**Plot histogram of 10,000 realizations of a Poisson distributed random variable.**

```{r}
poisson <- rpois(10000, 4.7)
hist(poisson)
```

**Using Iris dataset, write a for loop to calculate mean Sepal length for each species.**

```{r}
sp.l <- split(iris,iris$Species)
res <- list()
for(n in names(sp.l)){   
   dat <- sp.l[[n]] 
   res[[n]] <- data.frame(species=n,
                          mean.sepal.length=mean(dat$Sepal.Length))
                          
}
print(res)
```

**Using mtcars dataset, make scatter plot of mpg vs disp.**

```{r}
mtcarsplot <- ggplot(mtcars, aes(x = mpg, y = disp)) +
    geom_point(aes(color = factor(cyl))) +
stat_smooth(method = "lm",
        col = "#C42126",
        se = FALSE,
        size = 1)
linearmodel <- lm(disp ~ mpg, data = mtcars)
mtcarsplot
summary(linearmodel)
```

**Write a function to convert a vector of continuous data into its Z-scores.**

```{r}
z_score_function <- function(x) {
  z_score <- (x-mean(x))/sd(x)
  return(z_score)
}
vector_values=c(-4.89, -1.93, -1.11, 3.94, 0.46, -3.85, -0.20, 6.04,
9.36, 7.26, 4.88, 13.45, -2.93, 6.39, -16.22, -3.24, 6.86, 11.87, 1.81, -2.54)
z_score_function(vector_values)
```


