---
title: "Seamus_Gallivan_R_Final_Exam"
author: "Seamus Gallivan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook: default
---
#Seamus Gallivan#
#12/12/19#
#SPHU 4160 Final Exam#

#the tidyverse, dplyr, stringr, car and modeest package will be used for this R markdown file#
```{r}
library(tidyverse)
library(dplyr)
library(stringr)
library(modeest)
library(car)
```
# Compile date and session info#
```{r}
# print and get runtime ----
date()

# session info ----
sessionInfo()
```

#read in the nola_crime_2018 dataset#
```{r}
nola_crime <- read_csv(file="/Users/seamusgallivan/Documents/nola_crime_2018.csv")
```
#Question 1 and 2#
```{r}
str(nola_crime)
```
#Observations = 9970, Variables = 24#

#Question 3#
```{r}
length(unique(nola_crime$Item_Number))
```
#9384 unique item numbers#

#Question 4#
```{r}
nola_crime %>% 
  select(GNOCDC_LAB, Signal_Description, Item_Number) %>%
  filter(str_detect(Signal_Description, "AGG")) %>%
filter(GNOCDC_LAB=="IRISH CHANNEL")
```
#1 Aggravated Crime in The Irish Channel Neighorhood#

#Question 5#
```{r}
nola_crime <- nola_crime %>% 
  group_by(GNOCDC_LAB) %>%
  mutate(unique_item_numbers = n_distinct(Item_Number)) 
nola_crime %>%
  arrange(desc(unique_item_numbers)) %>%
  select(GNOCDC_LAB, unique_item_numbers) %>%
  distinct(GNOCDC_LAB, .keep_all = TRUE) 
```
#Two largest number of crime reports = FRENCH QUARTER & LITTLE WOODS#

#Question 6#
```{r}
nola_crime <- nola_crime %>% mutate(
  crime_rate = (unique_item_numbers / pop) * 1000
)
nola_crime %>%
  select(pop,unique_item_numbers,crime_rate,GNOCDC_LAB) %>%
  distinct(GNOCDC_LAB, .keep_all = TRUE) 
nola_crime %>%
  group_by(GNOCDC_LAB) %>%
  select(GNOCDC_LAB,crime_rate) %>%
  distinct(GNOCDC_LAB, crime_rate)
```

#Question 7#
```{r}
nola_crime %>%
  select(GNOCDC_LAB, pop) %>%
  arrange(desc(pop))
```
#Largest neighborhood by population = LITTLE WOODS, population = 44947#

#Question 8#
```{r}
nola_crime %>%
  select(pop, unique_item_numbers, GNOCDC_LAB, crime_rate) %>%
  arrange(crime_rate)
```
#Lowest crime rate = LAKEWOOD, population = 13989, unique crimes = 10#

#Question 9#
```{r}
nola_crime %>%
  filter(GNOCDC_LAB=="LAKEWOOD")
nola_crime <- nola_crime %>% 
  mutate(
    rate_ratio = crime_rate / 0.7148474
  )
nola_crime %>%
  select(GNOCDC_LAB, rate_ratio) %>%
  distinct(GNOCDC_LAB, rate_ratio)
```
#Question 10#
```{r}
nola_crime1 <- nola_crime %>%
  group_by(GNOCDC_LAB) %>%
  filter(str_detect(Signal_Description, "AGG")) %>%
  mutate(
    unique_agg_crimes = n_distinct(Item_Number)
  ) %>%
  mutate(
    unique_agg_crime_rate = (unique_agg_crimes / pop) * 1000
  ) 
nola_crime1 %>%
  group_by(GNOCDC_LAB) %>%
  select(GNOCDC_LAB, unique_agg_crime_rate) %>%
  filter(str_detect(GNOCDC_LAB, "FRENCH QUARTER"))
nola_crime1 <- nola_crime1 %>% 
  group_by(GNOCDC_LAB) %>%
  mutate(
    agg_rate_ratio = unique_agg_crime_rate / 0.9886633
  )
nola_crime1 %>%
  group_by(GNOCDC_LAB)%>%
  select(agg_rate_ratio, GNOCDC_LAB) %>%
  distinct(agg_rate_ratio, GNOCDC_LAB)
```
#Question 11#
```{r}
mfv(nola_crime$Signal_Description)
mfv(nola_crime$Charge_Description, na_rm = TRUE)
```
#most common cause for police report = "DISTURBANCE (DOMESTIC)"#
#most common charge that is filed = "THEFT"

#Question 12#
```{r}
nola_crime1 <- nola_crime %>%
  filter(Signal_Description == "DISTURBANCE (DOMESTIC)") %>%
  mutate(
    domestic_disturbances = n_distinct(Item_Number)
  )
nola_crime1 <- nola_crime1 %>%
  mutate(
    dom_dist_rate = (domestic_disturbances / pop) * 1000
  )
nola_crime1 %>%
  select(GNOCDC_LAB, dom_dist_rate) %>%
  distinct(GNOCDC_LAB, dom_dist_rate)
```
#Question 13#
```{r}
nola_crime2 <- nola_crime %>%
  group_by(GNOCDC_LAB,crime_rate) %>% 
  select(GNOCDC_LAB,crime_rate,Victim_Age) %>%
  summarize(m = mean(Victim_Age, na.rm = TRUE))
nola_crime2
  linearreg <- lm(m ~ crime_rate, data = nola_crime2)
  print(linearreg)
  summary(linearreg)
```
#Question 14#
```{r}
nola_crime %>%
  select(GNOCDC_LAB, crime_rate) %>%
  arrange(crime_rate)
```
#Since LAKEWOOD had the lowest crime rate, it also had the lowest police reponsiveness. LAKEWOOD had the least value of least amonunt of Signals per population#

#Question 15#
```{r}
poisson <- rpois(10000, 4.7)
hist(poisson)
```
#Question 16#
```{r}
sp.l <- split(iris,iris$Species)
res <- list()
for(n in names(sp.l)){   
   dat <- sp.l[[n]] 
   res[[n]] <- data.frame(species=n,
                          mean.sepal.length=mean(dat$Sepal.Length))
                          
}
print(res)
```
#Question 17#
```{r}
mtcarsplot <- ggplot(mtcars, aes(x = mpg, y = disp)) +
    geom_point(aes(color = factor(cyl))) +
stat_smooth(method = "lm",
        col = "#C42126",
        se = FALSE,
        size = 1)
linearmodel <- lm(disp ~ mpg, data = mtcars)
mtcarsplot
summary(linearmodel)
```
#Question 18#
```{r}
z_score_function <- function(x) {
  z_score <- (x-mean(x))/sd(x)
  return(z_score)
}
vector_values=c(-4.89, -1.93, -1.11, 3.94, 0.46, -3.85, -0.20, 6.04,
9.36, 7.26, 4.88, 13.45, -2.93, 6.39, -16.22, -3.24, 6.86, 11.87, 1.81, -2.54)
z_score_function(vector_values)
```
#EXTRA CREDIT#
#Question 20#
```{r}
install.packages(c("rgeos", "gpclib", "maptools", "sp"))
```
```{r}
library(raster)
library(sf)
library(rgeos)
library(maptools)
nola_map <- read_sf(dsn = "/Users/seamusgallivan/Downloads", layer = "Neighborhood_Statistical_Areas")
m <- merge(nola_crime, nola_map, by='GNOCDC_LAB')
p <- colorRampPalette(c("white", "red"))(128)
palette(p)
rate <- m$rate_ratio
cols <- (rate - min(rate))/diff(range(rate))*127+1
plot(nola_map, col=cols)
```

