---
title: "Seamus Gallivan R Take Home Assignment 4"
output: html_notebook
---
#your lines above produced the following errors: NA/NaN argument and object 'output' not found

#the tidyverse package will be required for this R markdown file#
#load necessary packages#
```{r}
library(tidyverse)
```
#include print runtime and session info as part of preamble
#read in necessary data#
```{r}
mydata <- read_csv(file="/Users/seamusgallivan/Downloads/ghana_bednet.csv")
```
#Question 1 and 2#
```{r}
str(mydata)
```
# number of observations = 26342
# number of variables = 8

#Question 3#
```{r}
mydata[mydata$sex=="F", "sex"] <- 1
mydata[mydata$sex==".", "sex"] <- NA
table1 <- table(mydata$sex)
prop.table(table1)
```
# proportion of children who are male = 0.5051061 or 50.51061%

#Question 4#
```{r}
mydata2 <- mydata[mydata$bednet==1 & mydata$sex==1,][1:8]
sum(mydata2$follyr, na.rm = TRUE)
```
# total number of years of follow-up experienced by females with intervention = 8395.184

#Question 5#
```{r}
mydata3 <- mydata[mydata$cluster==52,][1:8]
table2 <- table(mydata3$cluster, mydata3$outcome)
prop.table(table2)
```
# proportion of those who died in cluster 52 = 0.02432432 or 2.432432%

#Question 6#
# too may significant digits, round to nearest three or four
```{r}
lapply(mydata, mean, na.rm=TRUE)
lapply(mydata, sd, na.rm=TRUE)
```
# mean at age enrollment = 25.28693 months old
# standard deviation of age at enrollment = 18.02471 months old

#Question 7#
# your intermediate steps and approach are correct, but your calculations are off by one decimal point (0.2795 should be 0.02795)
```{r}
mydata4 <- mydata[mydata$outcome==1 & mydata$bednet==0,][1:8]
sum(mydata4$outcome, na.rm = TRUE)
mydata5 <- mydata[mydata$bednet==0,][1:8]
sum(mydata5$follyr, na.rm = TRUE)
mydata6 <- mydata[mydata$outcome==1 & mydata$bednet==1,][1:8]
sum(mydata6$outcome, na.rm = TRUE)
mydata7 <- mydata[mydata$bednet==1,][1:8]
sum(mydata7$follyr, na.rm = TRUE)

```
# Control
# 461/16492.83 = 0.2795 mortality rate
# Intervention 
# 396/16841.04 = 0.2351 mortality rate

#Question 8#
#this approach is fine but you neglected to label your axes; also, one of your clusters exceeds the limit you have on your y-axis, my suggestion would be to label the units (in this case proportion) on your y-axis to fit your data
```{r}
mydata8 <- mydata %>%
  group_by(cluster, outcome) %>%
  summarise(n = n()) %>%
  mutate(freq = n / sum(n))
mydata9 <- mydata8[mydata8$outcome==1,][1:4]
barplot(mydata9$freq)
```
#Question 9 & 10#
```{r}
reg <- glm(outcome~bednet,family = binomial(link = "logit"),data = mydata)
summary(reg)
```
# Being in the intervention cluster was associated with lower probability of a child dying #
# p-value = 0.01, coefficienct = -0.17937 #

#overall good. Be careful with doing hand calculations as it can easily introduce errors to your answer, you can run these operations on R to not encounter these errors; Your script worked through each problem in a correct manner; however, as a next step to improve your proficiency in R, I suggest utilizing piping operators more as they help reduce the multiple intermediate steps you used (I'm thinking more of your approach for question 7). 
